\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}
The genesis of this work traces back to February 2023, when initial discussions with Prof. Dr. Christin Seifert, M.Sc. Van Bach Nguyen, and Dr. Jörg Schlötterer laid the foundation for the research topic at hand. The spark was ignited from an attempt by M.Sc. Shreyasi Pathak to reproduce the findings of another paper, which, due to inherent randomness, yielded incomplete reproduction. Prof. Seifert presented this challenge, which we, through collective discussions, refined into a compelling Master’s thesis topic. The involvement of Prof. Dr.-Ing. habil. Carolin Birk post-April, brought a collaborative yet guiding light to the journey and constructive feedback throughout.\\

My profound gratitude to Prof. Dr. Christin Seifert, who ignited the spark for this research by offering me the topic and establishing the research environment. Her feedback significantly shaped the course of this thesis, with a substantial review towards the end that greatly contributed to the final shape of this work.\\

I am deeply thankful to Prof. Dr.-Ing. habil. Carolin Birk for her supervision and continuous support throughout this research journey. Her trust in my abilities provided the confidence and framework necessary for me to undertake this academic endeavor. Her insightful feedback and valuable contributions were instrumental in enhancing the quality and rigor of this work.\\

My sincere appreciation goes to M.Sc. Osman Alperen Koras, my daily supervisor, for his consistent guidance, support, and constructive feedback at every level of depth. Being the first point of contact for any issues regarding this thesis, his extensive feedback has been indispensable in navigating the intricacies of this research.\\

Special thanks are due to Dr. Jörg Schlötterer for his invaluable assistance in crafting the experimental design. His expertise and timely feedback ensured the robustness of our research methodology, thereby substantially contributing to the research design.\\

I am also grateful to M.Sc. Van Bach Nguyen for his significant contributions during the initial phases of the thesis, which were crucial in specializing the research problem and setting a firm foundation for the subsequent work.\\

Further appreciation goes to M.Sc. Shreyasi Pathak for her guidance on one of the critical aspects of this thesis. Her expertise and the baseline provided from her work played a key role in advancing the experiments and ensuring their success.\\

Furthermore, I extend my heartfelt gratitude to the Institute for AI in Medicine for providing access to their computer cluster, and to the entire team for their unwavering support and assistance throughout my time there. The experiments conducted for this thesis greatly benefited from their state-of-the-art hardware and facilities. Their guidance, friendship, and the collaborative environment they fostered have significantly enriched my academic experience.\\

I also extend my gratitude to Dr. Kemal Suntay from the Mathematics Department for his thorough review of this thesis. His feedback was instrumental in identifying areas for clarification, ensuring the work is comprehensible to readers with varying levels of familiarity with the subject matter. His external perspective enriched the exposition and contributed to the overall coherence of the thesis.\\

Lastly, I wish to express my thanks to the broader scientific community for their continuous efforts in advancing knowledge. Their dedication serves as an inspiration and has been the bedrock upon which this research stands.\\

\vspace{10mm}

\cleardoublepage{}

\clearpage
\thispagestyle{empty}
\null
\newpage

